---
title: "Interactive graph"
format: html
---

## Malnutrition among women - Underweight 18+ years (%) (2016)

<div id="plot-container">
</div>

<script src="https://cdn.jsdelivr.net/npm/d3@7"></script>
<script src="./scripts/myscript.js"></script>

<style>
* {
  font-family: sans-serif;
}
.tooltip {
  position: absolute;
  text
}
</style>


This interactive bubble chart visualizes the percentage of women aged 18 and older who are classified as underweight in various countries as of 2016. 

- Each **bubble** represents a different **country**.
- The **size of the bubble** corresponds to the **percentage** of underweight women, with larger bubbles indicating higher percentages.
- The **color** of the bubbles may represent different regions or groups, providing a visual distinction between countries.

The chart is interactive:

- Hovering over a bubble will display the **country's name** and the **exact percentage** of women who are underweight.

This visualization tool is essential for:

- **Policymakers**, to understand the scope of malnutrition in women for informed decision-making.
- **Healthcare professionals**, to identify regions where interventions might be most needed.
- **Researchers**, to study global health and nutrition trends.

By providing a clear and concise summary of complex data, the chart facilitates an accessible and engaging exploration of global health metrics.


```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE}

# Code to convert data to json for interactive graphs
if (!require("readr")) install.packages("readr")
if (!require("jsonlite")) install.packages("jsonlite")
if (!require("dplyr")) install.packages("dplyr")
library(readr)
library(jsonlite)
library(dplyr)
library(tidyverse)

# Load the data
nutrition_data <- read_csv("./data/Nutrition.csv")

# Preprocess and select the required column
nutrition_data_processed <- nutrition_data %>%
  mutate(Underweight_18_plus = as.numeric(replace(`Underweight 18+ years (%) (2016)`, `Underweight 18+ years (%) (2016)` == '–', NA))) %>%
  select(`Countries and areas`, Underweight_18_plus) %>%
  drop_na(Underweight_18_plus) %>%
  rename(country = `Countries and areas`, underweight_percentage = Underweight_18_plus)

# Convert the data to JSON
nutrition_json <- toJSON(nutrition_data_processed, pretty = TRUE)

# Write the JSON to a file
#write(nutrition_json, "./data/nutrition_data.json")

```


```{r, echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE}

# Code to convert data to json for interactive graphs
# Load necessary libraries
if (!require("readr")) install.packages("readr")
if (!require("jsonlite")) install.packages("jsonlite")
if (!require("dplyr")) install.packages("dplyr")
library(readr)
library(jsonlite)
library(dplyr)
library(tidyverse)

# Load the data
nutrition_data <- read_csv("./data/Nutrition.csv")

# Rename columns as per the new mapping
nutrition_data_renamed <- nutrition_data %>%
  rename(
    Countries = `Countries and areas`,
    `Weight at birth - Low Birthweight (2015)` = `Low birthweight\n (%)\n 2015`,
    `Weight at birth -  Unweighed at birth  (2015–2021)` = `Unweighed at birth\n (%)\n 2015–2021`,
    `Malnutrition among preschool-aged children (0–4 years of age)   Stunted (2020)` = `Stunted (%) \n (2020)`,
    `Malnutrition among preschool-aged children (0–4 years of age)   Wasted (2015-2022) Severe` = `Wasted\n (%)\n (2015-2022) - severe`,
    `Malnutrition among preschool-aged children (0–4 years of age)   Wasted (2015-2022) Moderate and Severe` = `Wasted\n (%)\n (2015-2022) - moderate and severe`,
    `Malnutrition among preschool-aged children (0–4 years of age)   Overweight (2020)` = `Overweight (%) \n (2020)`,
    `Vitamin A supplementation, (%) 2021` = `Vitamin A supplementation, \n full coveragea \n (6-59 months of age) (%) \n 2021`,
    `Malnutrition among school-aged children (5–19 years of age) 2016 Thinness` = `Thinness (%)`,
    `Malnutrition among school-aged children (5–19 years of age) 2016 Overweight` = `Overweight (%)`,
    `Malnutrition among women Underweight 18+ years (%) (2016)` = `Underweight 18+ years (%) (2016)`,
    `Malnutrition among women Anaemia 15–49 years (%) (2019)` = `Anaemia 15–49 years (%) (2019)`,
    `Percentage of households consuming iodized salt (2015–2021)` = `Percentage of households consuming iodized salt\n 2015–2021`
  )

# Convert the data to JSON
nutrition_json <- toJSON(nutrition_data_renamed, pretty = TRUE)

# Write the JSON to a file
#write(nutrition_json, "./data/nutrition_data_updated.json")

```
